Informing forest carbon inventories under the Paris Agreement using the Global Forest Carbon Database (ForC v4.0)

### 2023-06-09 doc/manuscript/ESSDmanuscript.Rmd
filling in/fixing references
includes temporary hack to fix labrière citation (manually changed citation key to labriere)

#39

Manual review of records was the limiting step for data submission to EFDB. 
We prioritized review of
(1) records from the Forest Global Earth Observatory [ForestGEO, @anderson-teixeira_ctfsforestgeo_2015; @davies_forestgeo_2021],
(2) studies with confidence intervals recorded in ForC (because uncertainty estimates are important to the IPCC),
(3) original publications containing large numbers of EFDB-relevant records,
and (4) records from tropical regions.
The latter criteria was motivated by the fact that although tropical forest is the single most important biome for climate change mitigation [@griscom_natural_2017;@griscom_national_2020], ground-based data on tropical forest C cycling tends to be more scarce due to a variety of challenges [@refs; @delima_making_2022], and tropical countries are more likely to apply Tier 1 methodology that bases forest C budgets on internationally defined IPCC default values [@romijn_assessing_2015].

@@

New data collection and analysis is needed to fill notable knowledge gaps.
While aboveground biomass stocks in particular have received -- and continue to receive -- by far the most research attention [Table `r table_variables`, @anderson-teixeira_carbon_2021;@dubayah_global_2020; @quegan_european_2019; @nisar_nasaisro_2018], production of an accurate global map of forest C stocks remains an ongoing challenge [@araza_decade_2023]. 
Other pools and variables remain very poorly quantified [Table `r table_variables`, @anderson-teixeira_carbon_2021], introducing substantive uncertainties into global forest C budgets [@pan_large_2011;@harris_global_2021].
Furthermore, data distribution is uneven across forest types and geographical regions (Figs. `r fig_map`-`r fig_histograms`).
For instance, data on C cycling of tropical forests -- particularly in Africa -- remains relatively sparse, in large part due to substantial barriers to data collection and distribution [@delima_making_2022].
Significant investment in research and researchers focused on ground-based measurement of forest C in such regions will be important to filling knowledge gaps in forest C cycling [@delima_making_2022;@araza_decade_2023; @labrière_forest_2023].

@@

Filling knowledge gaps in other EFDB-relevant variables will require more effort, but this effort is warranted given their importance for estimating forest C stock chnages.
Although aboveground biomass is the most studied variable considered here (Table `r table_variables`) and is the target of satellite missions [@dubayah_global_2020; @quegan_european_2019; @nisar_nasaisro_2018], significant ground-based research effort is required to create accurate global maps of forest biomass and changes therein [@duncanson_importance_2019; @labrière_forest_2023;@calders_laser_2022].
Given observations of increasing tree mortality in some forested regions [@mcdowell_pervasive_2020], better characterization of forest dead wood will be critical.
Additionally, C stocks in forest organic horizons and soils can be quite substantial and highly uncertain in many parts of the world [@tifafi_large_2018].
Significant investment in ground-based forest research will be critical to filling these gaps.

@@

\captionsetup[table]{labelformat=empty}
Table A1: **Table of changes to ForC fields.**
```{r table_ForCchanges, eval = TRUE, echo=FALSE, warning=FALSE, message = FALSE}
library (readr)
urlfile="https://raw.githubusercontent.com/forc-db/ForC/master/database_management_records/record_of_changes.csv"
record_of_changes <-read_csv(url(urlfile))
names(record_of_changes) <- stringr::str_replace_all(names(record_of_changes), "\\.", " ")
record_of_changes <- record_of_changes[c(1:5)] #include all fields but date of change
#record_of_changes<- record_of_changes[which(!grep("housekeeping",record_of_changes$Motivation)) , c(1:5)]  #trying to exclude rows for which Motivation = "housekeeping", but this is not working
record_of_changes[is.na(record_of_changes)]<- "" #replace NAs with blanks

>>>>

Manual review of records was the limiting step for data submission to EFDB. 
We prioritized review of
(1) records from the Forest Global Earth Observatory [ForestGEO, @anderson-teixeira_ctfsforestgeo_2015; @davies_forestgeo_2021],
(2) studies with confidence intervals recorded in ForC (because uncertainty estimates are important to the IPCC),
(3) original publications containing large numbers of EFDB-relevant records,
and (4) records from tropical regions.
The latter criteria was motivated by the fact that although tropical forest is the single most important biome for climate change mitigation [@griscom_natural_2017;@griscom_national_2020], ground-based data on tropical forest C cycling tends to be more scarce due to a variety of challenges [@anderson-teixeira_carbon_2021; @delima_making_2022], and tropical countries are more likely to apply Tier 1 methodology that bases forest C budgets on internationally defined IPCC default values [@romijn_assessing_2015].

@@

New data collection and analysis is needed to fill notable knowledge gaps.
While aboveground biomass stocks in particular have received -- and continue to receive -- by far the most research attention [Table `r table_variables`, @anderson-teixeira_carbon_2021;@dubayah_global_2020; @quegan_european_2019; @nisar_nasaisro_2018], production of an accurate global map of forest C stocks remains an ongoing challenge [@araza_decade_2023]. 
Other pools and variables remain very poorly quantified [Table `r table_variables`, @anderson-teixeira_carbon_2021], introducing substantive uncertainties into global forest C budgets [@pan_large_2011;@harris_global_2021].
Furthermore, data distribution is uneven across forest types and geographical regions (Figs. `r fig_map`-`r fig_histograms`).
For instance, data on C cycling of tropical forests -- particularly in Africa -- remains relatively sparse, in large part due to substantial barriers to data collection and distribution [@delima_making_2022].
Significant investment in research and researchers focused on ground-based measurement of forest C in such regions will be important to filling knowledge gaps in forest C cycling [@delima_making_2022;@araza_decade_2023; @labriere_forest_2023].

@@

Filling knowledge gaps in other EFDB-relevant variables will require more effort, but this effort is warranted given their importance for estimating forest C stock chnages.
Although aboveground biomass is the most studied variable considered here (Table `r table_variables`) and is the target of satellite missions [@dubayah_global_2020; @quegan_european_2019; @nisar_nasaisro_2018], significant ground-based research effort is required to create accurate global maps of forest biomass and changes therein [@duncanson_importance_2019; @labriere_forest_2023;@calders_laser_2022].
Given observations of increasing tree mortality in some forested regions [@mcdowell_pervasive_2020], better characterization of forest dead wood will be critical.
Additionally, C stocks in forest organic horizons and soils can be quite substantial and highly uncertain in many parts of the world [@tifafi_large_2018].
Significant investment in ground-based forest research will be critical to filling these gaps.

@@

\captionsetup[table]{labelformat=empty}
Table A1: **Table of changes to ForC fields.** These are changes implemented between releases of v3.0 and v4.0. 
```{r table_ForCchanges, eval = TRUE, echo=FALSE, warning=FALSE, message = FALSE}
library (readr)
urlfile="https://raw.githubusercontent.com/forc-db/ForC/master/database_management_records/record_of_changes.csv"
record_of_changes <-read_csv(url(urlfile))
names(record_of_changes) <- stringr::str_replace_all(names(record_of_changes), "\\.", " ")
record_of_changes <- record_of_changes[c(1:5)] #include all fields but date of change
#record_of_changes<- record_of_changes[which(!grep("housekeeping",record_of_changes$Motivation)) , c(1:5)]  #trying to exclude rows for which Motivation = "housekeeping", but this is not working
record_of_changes[is.na(record_of_changes)]<- "" #replace NAs with blanks

### 2023-06-09 doc/manuscript/ESSDmanuscript.tex
filling in/fixing references
includes temporary hack to fix labrière citation (manually changed citation key to labriere)

#39

Manual review of records was the limiting step for data submission to
EFDB. We prioritized review of (1) records from the Forest Global Earth
Observatory
\citep[ForestGEO,][]{anderson-teixeira_ctfsforestgeo_2015, davies_forestgeo_2021},
(2) studies with confidence intervals recorded in ForC (because
uncertainty estimates are important to the IPCC), (3) original
publications containing large numbers of EFDB-relevant records, and (4)
records from tropical regions. The latter criteria was motivated by the
fact that although tropical forest is the single most important biome
for climate change mitigation
\citep{griscom_natural_2017, griscom_national_2020}, ground-based data
on tropical forest C cycling tends to be more scarce due to a variety of
challenges \citep{refs, delima_making_2022}, and tropical countries are
more likely to apply Tier 1 methodology that bases forest C budgets on
internationally defined IPCC default values
\citep{romijn_assessing_2015}.

@@

As of June 09, 2023, we had reviewed or added 2292 EFDB-relevant
records, 1438 records of which were submitted to EFDB, and 376 of which
have been reviewed, accepted, and posted (Figs. 2-3, Table 2). The 37\%
attenuation between records reviewed and those sent to EFDB was
attributable to the presence of digitized records and records where a
variable's value had been calculated as the sum or difference of related
variables rather than presented directly in the text. The discrepancy
between the number of records sent and that posted to EFDB is primarily
attributable to the time required for the IPCC to review the records,
and it is not expected that many -- if any -- records will be rejected.

@@

New data collection and analysis is needed to fill notable knowledge
gaps. While aboveground biomass stocks in particular have received --
and continue to receive -- by far the most research attention
\citep[Table
2,][]{anderson-teixeira_carbon_2021, dubayah_global_2020, quegan_european_2019, nisar_nasaisro_2018},
production of an accurate global map of forest C stocks remains an
ongoing challenge \citep{araza_decade_2023}. Other pools and variables
remain very poorly quantified \citep[Table
2,][]{anderson-teixeira_carbon_2021}, introducing substantive
uncertainties into global forest C budgets
\citep{pan_large_2011, harris_global_2021}. Furthermore, data
distribution is uneven across forest types and geographical regions
(Figs. 2-3). For instance, data on C cycling of tropical forests --
particularly in Africa -- remains relatively sparse, in large part due
to substantial barriers to data collection and distribution
\citep{delima_making_2022}. Significant investment in research and
researchers focused on ground-based measurement of forest C in such
regions will be important to filling knowledge gaps in forest C cycling
\citep{delima_making_2022, araza_decade_2023, labrière_forest_2023}.

@@

Filling knowledge gaps in other EFDB-relevant variables will require
more effort, but this effort is warranted given their importance for
estimating forest C stock chnages. Although aboveground biomass is the
most studied variable considered here (Table 2) and is the target of
satellite missions
\citep{dubayah_global_2020, quegan_european_2019, nisar_nasaisro_2018},
significant ground-based research effort is required to create accurate
global maps of forest biomass and changes therein
\citep{duncanson_importance_2019, labrière_forest_2023, calders_laser_2022}.
Given observations of increasing tree mortality in some forested regions
\citep{mcdowell_pervasive_2020}, better characterization of forest dead
wood will be critical. Additionally, C stocks in forest organic horizons
and soils can be quite substantial and highly uncertain in many parts of
the world \citep{tifafi_large_2018}. Significant investment in
ground-based forest research will be critical to filling these gaps.

@@

Table A1: \textbf{Table of changes to ForC fields.}
\begingroup\fontsize{8}{10}\selectfont

>>>>

Manual review of records was the limiting step for data submission to
EFDB. We prioritized review of (1) records from the Forest Global Earth
Observatory
\citep[ForestGEO,][]{anderson-teixeira_ctfsforestgeo_2015, davies_forestgeo_2021},
(2) studies with confidence intervals recorded in ForC (because
uncertainty estimates are important to the IPCC), (3) original
publications containing large numbers of EFDB-relevant records, and (4)
records from tropical regions. The latter criteria was motivated by the
fact that although tropical forest is the single most important biome
for climate change mitigation
\citep{griscom_natural_2017, griscom_national_2020}, ground-based data
on tropical forest C cycling tends to be more scarce due to a variety of
challenges \citep{anderson-teixeira_carbon_2021, delima_making_2022},
and tropical countries are more likely to apply Tier 1 methodology that
bases forest C budgets on internationally defined IPCC default values
\citep{romijn_assessing_2015}.

@@

As of June 09, 2023, we had reviewed or added 2292 EFDB-relevant
records, 1438 records of which were submitted to EFDB, and 376 of which
have been reviewed, accepted, and posted (Figs. 2-3, Table 2). The 37\%
attenuation between records reviewed and those sent to EFDB was
attributable to the presence of digitized records and records where a
variable's value had been calculated as the sum or difference of related
variables rather than presented directly in the text. The discrepancy
between the number of records sent and that posted to EFDB is primarily
attributable to the time required for the IPCC to review and post the
records, and also because a minority of records were deemed not
applicable to EFDB by the review panel.

@@

New data collection and analysis is needed to fill notable knowledge
gaps. While aboveground biomass stocks in particular have received --
and continue to receive -- by far the most research attention
\citep[Table
2,][]{anderson-teixeira_carbon_2021, dubayah_global_2020, quegan_european_2019, nisar_nasaisro_2018},
production of an accurate global map of forest C stocks remains an
ongoing challenge \citep{araza_decade_2023}. Other pools and variables
remain very poorly quantified \citep[Table
2,][]{anderson-teixeira_carbon_2021}, introducing substantive
uncertainties into global forest C budgets
\citep{pan_large_2011, harris_global_2021}. Furthermore, data
distribution is uneven across forest types and geographical regions
(Figs. 2-3). For instance, data on C cycling of tropical forests --
particularly in Africa -- remains relatively sparse, in large part due
to substantial barriers to data collection and distribution
\citep{delima_making_2022}. Significant investment in research and
researchers focused on ground-based measurement of forest C in such
regions will be important to filling knowledge gaps in forest C cycling
\citep{delima_making_2022, araza_decade_2023, labriere_forest_2023}.

@@

Filling knowledge gaps in other EFDB-relevant variables will require
more effort, but this effort is warranted given their importance for
estimating forest C stock chnages. Although aboveground biomass is the
most studied variable considered here (Table 2) and is the target of
satellite missions
\citep{dubayah_global_2020, quegan_european_2019, nisar_nasaisro_2018},
significant ground-based research effort is required to create accurate
global maps of forest biomass and changes therein
\citep{duncanson_importance_2019, labriere_forest_2023, calders_laser_2022}.
Given observations of increasing tree mortality in some forested regions
\citep{mcdowell_pervasive_2020}, better characterization of forest dead
wood will be critical. Additionally, C stocks in forest organic horizons
and soils can be quite substantial and highly uncertain in many parts of
the world \citep{tifafi_large_2018}. Significant investment in
ground-based forest research will be critical to filling these gaps.

@@

Table A1: \textbf{Table of changes to ForC fields.} These are changes
implemented between releases of v3.0 and v4.0.
\begingroup\fontsize{8}{10}\selectfont
